---
title: "Assignment 1"
author: "Dheeraj Singh"
date: "January 6, 2018"
output: html_document
---
# Reproducible Research Peer Graded Assignment- 1

## Add all the required packages in the workspace

```{r, message=FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(lubridate)
library(stringr)

```

## Read file code of "Activity Monitoring Data"
```{r}
amd <- fread("activity.csv", header = TRUE)
```

## Have a look on the summary
```{r}
head(amd, n = 10)
summary(amd)
glimpse(amd)
```

```{r}
## Change the date format whcih is saved as a charecter vecrtor

amd$date <- ymd(amd$date)
```


## Arrange data set as per requirement and **Histogram of TOtal no of step taken per day**
```{r}
amd %>% filter(!is.na(steps))%>%
        group_by(date)%>%
        summarize(spd = sum(steps))-> amd1

amd %>% filter(!is.na(steps))%>%
        group_by(date)%>%
        summarize(spd = sum(steps))-> amd1

ggplot(amd1, aes(x = spd))+
        geom_histogram(binwidth = 1000, col  ="red")+
        labs(title = "Histogram for Steps per Day", x = "Steps", y= "Count of steps")
```


## Calculating mean and median of each day of dataset
```{r}
step_mean <- mean(amd1$spd, na.rm = T)
step_median <- median(amd1$spd, na.rm = T)
step_mean
step_median
```


## Time Series plot to make a chart

```{r}

amd %>% filter(!is.na(steps))%>%
group_by(interval)%>%
        summarize(steps = mean(steps))-> amd1

ggplot(amd1, aes(x = interval, y = steps, type = "l"))+
        geom_line(col  ="red")+
        labs(title = "Time-Series Plot for Steps", x = "Interval", y=  "Steps")
```
             
## Code to describe and show a strategy for imputing missing data

to impute missing values we would replace all NA's with mean of steps'

```{r}
amd$steps[is.na(amd$steps)] <- mean(amd$steps, na.rm = T)
head(amd)
sum(is.na(amd$steps))
```


## Histogram of the total number of steps taken each day after missing values are imputed
```{r}
amd %>% group_by(interval)%>%
        summarize(steps = mean(steps))-> amd2

ggplot(amd2, aes(x = interval, y = steps, type = "l"))+
        geom_line(col  ="red")+
        labs(title = "Time-Series Plot for Steps", x = "Interval", y=  "Steps")
```


## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

First we need to mutate a new coloumn of weektype and then plotting the graph of comparision of weekdays and weekedends no of steps

```{r}
amd_full <- amd %>% mutate(weektype = ifelse(weekdays(date) == "Saturday" | weekdays(date) == "Sunday", "Weekend", "Weekdays"))
amd_full %>% group_by(interval, weektype)%>%
        summarize(stepmean = mean(steps))%>%
        ggplot(aes(x = interval, stepmean))+
        geom_line(col = "red")+
        facet_wrap(~weektype, ncol = 1, nrow=2)+
        labs(x = "interval", y = "Average no of steps")
```
